\documentclass[letterpaper,12pt]{exam}

\RequirePackage{comment}
\RequirePackage{GE05}
% this inputs graphicx, too
\RequirePackage[hypertex]{hyperref}
    \hypersetup{colorlinks=true,urlcolor=blue,linkcolor=red}
    
\newcommand{\POP}{\mbox{\it POP\/}}

\def\ClassName{The Global Economy}
\def\Category{Professor David Backus}
\def\HeadName{Group Project \#3}

\printanswers 

\begin{document}
\parindent = 0.0in
\parskip = \bigskipamount
\thispagestyle{empty}%
\Head

\centerline{\large \bf \HeadName:  Sources of Country Performance}%
\centerline{Revised:  \today}

\medskip
{\it Submit via Blackboard by 9am February 16.}

\begin{questions}
%==============================================================================
\question {\it China in 2030.} 
China's remarkable economic growth and large
population have created one of the world's largest economies. In
2004, GDP in China was slightly greater than half the size of US
GDP, but growing twice as fast. The question is how this is likely
to change in the coming years. 
Will China become the world's largest economy?  

Your mission is to estimate GDP for the US and China in 2030 
using the Solow model and some inspired estimates of
the parameters. In 2004, the economies had the following
characteristics:
%
\begin{center}
\tabcolsep = 0.17in
\begin{tabular}{lcccc}
\hline\hline
Country   &  GDP &  Capital &  Employment &  Population \\
\hline\hline
China   &  5,592 &  22,276   & 747.36  & 1,286.98\phantom{1,} \\%
US      & 10,761 &  31,672   & 141.93  &  292.62 \\%
\hline\hline
\end{tabular}
\end{center}
%
GDP and capital are reported in billions of 2000 US dollars,
employment and population in millions.  Your economic consultants
tell you that the parameter values are
%
\begin{center}
%\tabcolsep = 0.1in
\begin{tabular}{lcccc}
\hline\hline
Country \hspace{0.1in}  &  Saving &  Depreciation &  Empl. Growth &  TFP Growth    \\
\hline\hline
China   &  35\% &  6\%   &   1.0\%  & 4\%  \\%
US      &  20\% &  6\%   &   0.5\%  & 2\%  \\%
\hline\hline
\end{tabular}
\end{center}
%
Data sources. GDP and Employment are from the World Bank. Capital is
our estimate, based on data on investment from the PWT and the World
Bank.
%
\begin{parts}
\part Compute TFP for each country in 2004 using the basic
production function, $Y = A K^\alpha L^{1-\alpha}$. (5~points)

\part Compute and graph time paths for GDP in both countries 
starting in 2004 and ending in 2030. 
Which economy is larger (in GDP terms) in 2030?  
(25~points)

\part Which parameter inputs are most important to your answer?  
What happens, for example, if China's saving rate falls to 25\%?  
TFP growth falls to 3\%?  
(10~points) 

\part How do GDP per capita in China and the US compare in 2030?
How might this difference influence your approach to (say) consumer 
products?  
(10~points) 

\end{parts} 

\begin{solution}
\begin{parts}
\part The usual calculation:  $A = Y /(K^{\alpha}
L^{1-\alpha})$.  
The numbers are 2.41 for China and 12.50 for the US.

\part This is a dynamic simulation much like the one described in
``Solow's Model of Economic Growth.''  
I did it using the spreadsheet we set up in class.  

We simulate the model like this.  At each date, we compute
employment $L$ and productivity $A$ from their values at the
previous date by using their (assumed constant) growth rates:
\begin{eqnarray*}
    L_{t+1} &=& (1+n) L_t \\
    A_{t+1} &=& (1+a) A_t,
\end{eqnarray*}
where $n$ and $a$ are the growth rates of employment and TFP,
respectively. Capital evolves as in the Solow model:
\begin{eqnarray*}
    K_{t+1} &=& s A_t K_t^\alpha L_t^{1-\alpha} + (1 - \delta) K_t .
\end{eqnarray*}
All we need are the parameters, which are given to us in the
problem. 
The solution includes \\

\begin{center}
\tabcolsep = 0.2in
\begin{tabular}{lrr}
\hline\hline Year   &  China   &  US  \\
\hline\hline
    2004  &  5,592   &  10,761  \\
    2005  &  5,908   &  11,042  \\
    2006  &  6,247   &  11,335  \\
    2007  &  6,611   &  11,640  \\
    2008  &  7,001   &  11,958  \\
    2009  &  7,419   &  12,289  \\
    2010  &  7,868   &  12,634  \\
    2015  &  10,648  &  14,579  \\
    2020  &  14,592  &  16,943  \\
    2030  &  28,084  &  23,243  \\
    2050  &  108,813 &  45,303  \\
\hline\hline
\end{tabular}
\end{center}
\bigskip

In this scenario, China overcomes about half the difference 
between the two countries by 2020.  
If you keep running the model, China surpasses the US in 2025.  


\part
This seems a lot like extrapolating a current trend, which is a 
notoriously bad way to predict the future.  
Trends change and extrapolation, by design, cannot predict changes.  
That's not quite the case, because the model makes it clear what 
we need to come up with the answer --- primarily, 
the parameter values.  
All of them can and should be questioned.  
One is the rate of TFP growth:  will it fall 
as China approaches the world's
``technology frontier''?
Another is the investment rate:  will it fall as living standards improve?
A good analysis would include some discussion of these issues
and their consequences for GDP.  

\part 
In 2004, GDP per worker in the US is about 10 times that of China.
By 2030, the ratio is about 5.  
This is a huge change, but the difference is still large. 
If we try to compute GDP per capita, as requested, 
we face a data issue:  what is population is 2030?  
This is the kind of annoying little data issue that comes up all the time
in real life.
Consulting interviews, for example, hit you with this kind of thing.
The key is to not let it get in our way.  
One thing we could do is guess that the ratio of employment to population
doesn't change.  
In that case we find that GDP per capita in 2030 
is about 70,000 in the US and 17,000 in China, 
making the average consumer in the US about 4 times richer
than the average consumer in China.  

A more sophisticated answer would be to look up population projections
and plug them into your analysis.  
If you did that, you get some credit for initiative.
If you want to push this further, you might ask how 
the population and age distribution might feed back on
employment rates.  
Do the population projections seem consistent with our 
assumption about employment growth?  

How would tall this affect your approach to consumer products?
The income distribution matters, too, 
but on the whole you're likely to be selling less
expensive products in China.  
The big international consumer products companies all 
have a presence in China and that's one of the things
they've found.
Over time, the differences will shrink, but they have a 
ways to go.  
\end{parts}
\end{solution} 


%=============================================================================
\question {\it Argentina and Chile. } 
Argentina and Chile have each experienced 
both dramatic growth and painful reversals over the last century.  
Argentina was one of the richest countries in the world in 1900, 
but mixed economic performance since then has 
dropped it back into the world of emerging markets.  
Chile suffered a traumatic change in government in the 1970s, 
but emerged a decade later as a fast-growing democracy 
and now has the highest per capita GDP in Latin America.  
Your mission is to tell us how --- and maybe why 
--- their experiences have been different.  

\begin{center}
%\tabcolsep = 0.12in
\begin{tabular}{lccccc}
\hline\hline
Country   &  Year   &  POP   &  Employment &  Y/POP &  K/POP  \\
\hline\hline
Argentina \hspace*{0.1in} &  1960   &  20.6m &  \phantom{1}8.1m   
                &  \phantom{1}7,838  &  12,713  \\
Argentina &  2004   &  39.1m &  16.2m             &  10,939 &  24,343  \\
Chile   &  1960  & 7.59m &  2.55m   
                &  \phantom{1}5,086   &   16,666  \\
Chile     &  2004   & 15.67m\phantom{1} &  6.57m   &  12,678  &  29,437  \\
\hline\hline
\end{tabular}
\end{center}
POP is population (millions), Employment is the number of people working 
(millions), Y is GDP (2000 US dollars), 
and K is capital (2000 US dollars).  
Y and K are PPP-adjusted.  
Data are from the Penn World Tables.  

\begin{parts} 
\part Compute the (average annual continuously compounded) 
growth rates of GDP per capita and GDP per worker in the two countries.  (15~points)

\part Use our growth accounting methodology to allocate growth in
output per worker to TFP and capital per worker.
Which factor has been most important over the last 40+ years? (25~points)

%\item Institutions??  

\part Using whatever resources you think appropriate, 
comment on the differences in performance in the two countries 
and their likely performance over the next decade.  
Which do you think grow faster?  
(10~points) 
\end{parts}


\begin{solution}
GDP per capita in Chile has gone 
from 35\% below Argentina in 1960 to 16\% above in 2004.
How did this happen?
Most of this problem is concerned with aggregate data:
how much comes from capital, TFP, etc?  
The last part is an opportunity to look further 
into features of the two economies that might account 
for what we find in the data analysis.  
%
\begin{parts}
\item The (continuously compounded annual) 
growth rate of GDP per capita for Argentina is 
\[
    \gamma \;=\; \log (10,939/7,838) / (2004 - 1960) \;=\; 0.76\%. \; (!) 
\]
Refer to the discussion of growth rates in ``Sources of Growth''
if you're not sure why this works.
A similar calculation gives us 2.08\% for Chile.  
GDP per worker is GDP per capita times the population then
divided by the number of workers.  
Its growth rates are 0.64\% (Argentina) and 1.57\% (Chile).  
Bottom line:  faster growth in Chile.  

\part We use the production function 
\[
    Y/L \;=\; A (K/L)^{1/3} ,
\]
so TFP is $ A = (Y/L) / (K/L)^{1/3}  $. 
In growth rates, with the adjustment for employment, 
this translates into 
\begin{eqnarray*}
    \gamma_{Y/POP} &=&  \gamma_{L/POP}  + \gamma_A 
                + \alpha \gamma_{K/L}  ,
\end{eqnarray*}
where $\alpha = 1/3$ as usual.  
This allows us to decompose growth in per capita GDP into 
components due to the employment rate, productivity, 
and capital per worker.  
A few calculations give us 
%
\begin{center}
%\tabcolsep = 0.12in
\begin{tabular}{lcccc}
\hline\hline
            &  $Y/\POP$      &  $L/\POP$    &   $A$   &  $K/L$     \\
\hline\hline
\multicolumn{5}{l}{\it Argentina} \\
1960        &    7,838    &  0.393  &   626  &  32,332      \\
2004        &    10,939\phantom{1}   
                          &  0.414  &   679  &  58,754    \\
Growth rate (\%) &  0.76  &  0.12   &   0.19 &  1.36   \\
Contribution to growth (\%) 
                &   0.76  &  0.12   &   0.19 &  0.45  \\
\hline                 
\multicolumn{5}{l}{\it Chile} \\
1960        &    5,086    &  0.336   &  412  & 49,616      \\
2004        &    12,678\phantom{1}
                          &  0.419   &  733  & 70,210    \\
Growth rate (\%) &  2.08  &  0.50   &   1.31 &   0.79   \\
Contribution to growth (\%) 
                &   2.08  &  0.50   &  1.31  &   0.26  \\
\hline\hline
\end{tabular}
\end{center}
%
Here ``growth rate'' is computed as in (a) and multiplied 
by 100 to give us a percentage.
``Contribution'' modifies these growth
rates as they occur in the formula:  the growth rate of $K/L$ gets
multiplied by $\alpha = 1/3$ (its exponent in the production function).
The numbers tell us that almost all of the difference in growth 
comes from TFP.  

The question calls for a slightly simpler calculation:
to compute the components in growth of GDP per worker.  
It's the same, except we skip the growth rate of $L/\POP$.
In Argentina, for example, the growth rate of GDP per worker (0.64\%)
consists of growth in capital per worker (0.45\%) 
and growth in TFP (0.19\%).  

\part Your call, but here are some things you could mention.  
Almost all of the institutional indicators are stronger for Chile
than Argentina, so perhaps it's not surprising that economic performance
has been better.  
Much of this has come in the last 25 years, as Chile has reestablished a 
stable democracy with sensible economic policies.
Argentina continues to fluctuate between policies that favor long-term 
performance and those that focus on the short term.  

My opinion is that the strong economic and political 
foundations Chile has laid will continue to generate higher growth 
than we'll see in Argentina. 
For one thing, we've had different parties in power over the last 20 years, 
yet the same basic economic policies are followed.  
One that I find intriguing is a privatized social security system, 
which helped build a domestic capital market.  
A temporary factor is copper prices, which are unusually high right now. 
It's a tricky measurement issue, but this tends to show up as high GDP 
and therefore high TFP.  
History tells us that probably won't last.
With that in mind, the government saved revenue from copper 
for several years, 
despite intense political pressure to spend it.  
When the crisis hit, they had a pot of money they could use to mitigate
its effects. 

\end{parts}
\end{solution}

\end{questions} 

\vfill \centerline{\it \copyright \ \number\year \  
NYU Stern School of Business}

\end{document}

\begin{enumerate}

\item {\it Japan and Korea. }   
Japan and Korea are remarkable success stories, 
having experienced unusually rapid economic growth over long periods of time.
In Japan, per capita GDP rose from 30\% of the US in 1870 
(similar to Mexico today) to about 75\% today.  
In Korea, GDP per capita rose from about 10\% of the US in 1960 
(similar to the Philippines at the time) 
to over 50\% today.
Your mission:  to determine the sources of the remaining differences
between the two countries --- capital, productivity, and anything else 
that you think is relevant.  

The data for 2003 
(the most recent year for which all of these numbers are available):
\begin{center}
%\tabcolsep = 0.12in
\begin{tabular}{lcccccc}
\hline\hline
        &  POP  &  Y/POP  &  L/POP  &  K/Y     &  Education &  Hours \\
\hline\hline
Japan   & 127.7m &  24,036 &  0.534 & 4.003     & \phantom{1}9.6 & 1801 \\%
Korea   &  \phantom{1}47.5m 
                 &  17,600 &  0.494 & 1.608    & 11.0  &  2434  \\%
\hline\hline
\end{tabular}
\end{center}
POP is population (millions), Y is GDP (2000 US dollars), 
K is capital (2000 US dollars), Education is years of school, 
and Hours is annual hours worked.  
Y and K are PPP-adjusted.  
Education and Hours are from the OECD's {\it Employment Outlook\/};
the other variables are from the Penn World Tables.  


\begin{enumerate}
\item Compute output per capita, output per worker, and output per
hour worked.  How do they differ? (15~points)

\item What are the primary sources of the difference in output per
capita? (15~points)

\item Using whatever resources you think appropriate, 
comment on the likely performance of the two countries over the next 
decade.  
What do you see as the major opportunities?  Challenges?  (20~points) 
\end{enumerate}


%\newpage
%\begin{comment}
Answer.
\begin{enumerate}
\item As we know from class, GDP per capita, per worker, and per
hour worked are simply the ratios of GDP to population, number of
workers, and total hours worked, respectively. The numbers are:
\begin{center}
\begin{tabular}{lccc}
\hline\hline
Country   &  GDP Per Capita &  GDP Per Worker &  GDP Per Hour Worked \\
\hline\hline
France  &  26.673  & 65.01 &  0.0447 \\
US      &  35.356  & 73.19 &  0.0408 \\
France/US &  0.754 &  0.888 &  1.095 \\
\hline\hline
\end{tabular}
\end{center}
Note that the ratios (the bottom line) increase as we move to the
right:  the difference in GDP per person is due to differences in the amount of work (workers and hours).

\item A natural version of the production function in this case is
\[
    Y \;=\; A K^\alpha (H h L)^{1-\alpha} ,
\]
where $h$ is average hours worked and $L$ is the number of workers
(employment).  We find $A$ by substituting in for all the other
numbers:  $ A =Y / [ K^\alpha (H h L)^{1-\alpha}] $.  How do we
use this to make sense of output per capita? If POP is the
population, then
\[
    Y/\POP \;=\; (Y/L) (L/\POP) \;=\; (L/\POP) A (K/L)^\alpha (H
    h)^{1-\alpha} .
\]
?? ** do more on education ??
This isn't the only possible line of attack on this problem, but
it's a reasonable one.  It tells us we can attribute differences
in per capita GDP to differences in:  participation ($L/\POP$),
total factor productivity ($A$), capital per worker ($K/L$), human
capital ($H$), and average hours per worker ($h$). The math
implies
%
\begin{eqnarray*}
    \frac{(Y/\POP)_{F}}{(Y/\POP)_{US}} &=&
             \left[ \frac{(L/\POP)_F}{(L/\POP)_{US}} \right]
             \left[ \frac{A_F}{A_{US}} \right]
             \left[ \frac{(K/L)_{F}}{(K/L)_{US}} \right]^\alpha
             \left[ \frac{H_{F}}{H_{US}} \right]^{1-\alpha}
             \left[ \frac{h_{F}}{h_{US}} \right]^{1-\alpha} \\
             &=& \left[ \frac{.410}{.483} \right]
                 \left[ \frac{.0158}{.0147} \right]
                 \left[ \frac{232.7}{201.3} \right]^{1/3}
                 \left[ \frac{11.9}{13.8} \right]^{2/3}
                 \left[ \frac{1453}{1792} \right]^{2/3}_{\phantom{\sum_i}} \\
             &=& (0.849)(1.074) (1.050) (0.862) (0.870) \;=\;
             0.754 .
\end{eqnarray*}
%
Where does the 25\% difference in GDP per person come from?
In order of importance:  ratio of employment to population, education, 
and hours worked.  
Capital per worker and TFP go the other way ---
they're higher in France. Obviously there are many caveats to this
analysis.  A major one is about education, which we have measured
with average numbers of completed years of schooling for the
population older than 15. Obviously, this is not the only way. For
example, we could try to correct this indicator for the quality of
education. France boasts the lowest number of pupils per teacher
in the world. If we accounted for this, the differences in human
capital between the two countries would not be as stark.
Another is that a lower employment rate may lead to a higher 
average skill level, with the least skilled not working.  

\item Evidently labor market outcomes are very different in the
two countries. Our conjecture is that differences in the
institutions and legislation that regulates the functioning of
labor markets play a major role. The US is probably a better place
to work, at least if you want to work long hours.  
We still love France as a place to live and eat, though.

\end{enumerate}
%\end{comment}

%\end{enumerate} \end{document}


\end{enumerate}
%\end{comment} 

\end{enumerate}


\vfill \centerline{\it \copyright \ \number\year \  NYU Stern
School of Business}


\end{document}
\item {\it Japan and Korea. }   
Japan and Korea are two of the most remarkable economic 
success stories in human history.
Having experienced unusually rapid economic growth over long periods of time, they have moved from the ranks of developed countries to 
developed and middle income countries, respectively.  
In Japan, per capita GDP rose from 30\% of the US in 1870 
(similar to Mexico today) to about 75\% today.  
In Korea, GDP per capita rose from about 10\% of the US in 1960 
(similar to the Philippines at the time) 
to over 50\% today.
Your mission:  to determine the sources of the remaining differences
between the two countries --- capital, productivity, and anything else 
that you think is relevant.  

The data for 2003 
(the most recent year for which all of these numbers are available):
\begin{center}
%\tabcolsep = 0.12in
\begin{tabular}{lcccccc}
\hline\hline
        &  POP  &  Y/POP  &  L/POP  &  K/Y     &  Education &  Hours \\
\hline\hline
Japan   & 127.7m &  24,036 &  0.534 & 4.003     & \phantom{1}9.6 & 1801 \\%
Korea   &  \phantom{1}47.5m 
                 &  17,600 &  0.494 & 3.366    & 11.0  &  2434  \\%
\hline\hline
\end{tabular}
\end{center}
POP is population (millions), Y is GDP (2000 US dollars), 
K is capital (2000 US dollars), Education is years of school, 
and Hours is annual hours worked.  
Y and K are PPP-adjusted.  
Education and Hours are from the OECD's {\it Employment Outlook\/};
the other variables are from the Penn World Tables.  

\begin{enumerate}
\item Compute output per capita, output per worker, and output per
hour worked.  How do they differ? (20~points)

\item What are the primary sources of the difference in output per
capita? (20~points)

\item Using whatever resources you think appropriate, 
comment on the likely performance of the two countries over the next 
decade.  
What do you see as the major opportunities?  Challenges?  (10~points) 
\end{enumerate}


%\newpage
%\begin{comment}

Answer.
Per capita output is higher in Japan than Korea, but why?
Is the Japanese economy more productive?  Other reasons?  
What would it take for Korea to catch up?  
\begin{enumerate}
\item Calculations are reported below.  
The most striking difference in the numbers:  
a large difference in average hours worked means 
that Japan's 26\% advantage in output per worker 
increases to 71\% in output per hour.  
In words:  Korea's gap in GDP per worker would be substantially 
larger if they didn't work so much.  
%
\begin{center}
\begin{tabular}{lccc}
\hline\hline
Country   &  GDP Per Capita &  GDP Per Worker &  GDP Per Hour \\
\hline\hline
Japan       &  24,036  & 45,011 &  24.99 \\
Korea       &  17,600  & 35,628 &  14.64 \\
Japan/Korea &  1.37 &  1.26 &  1.71 \\
\hline\hline
\end{tabular}
\end{center}

\item The idea is to dig a little deeper, 
introducing (for example) capital and productivity.  
You could do this in a variety of ways; 
one is to use 
\[
    Y \;=\; A K^\alpha (h L)^{1-\alpha} 
\]
to account for the difference in output per hour worked.  
A little multiplying and dividing gives us 
\[
    Y/\POP \;=\; (Y/L) (L/\POP) 
            \;=\; (L/\POP) A (K/L)^\alpha h^{1-\alpha} .
\]
This isn't the only possible line of attack on this problem, but
it's a reasonable one.  It tells us we can attribute differences
in per capita GDP to differences in:  participation ($L/\POP$),
total factor productivity ($A$), capital per worker ($K/L$), 
and average hours per worker ($h$). 
The math implies
%
\begin{eqnarray*}
    \frac{(Y/\POP)_{J}}{(Y/\POP)_{K}} &=&
             \left[ \frac{(L/\POP)_J}{(L/\POP)_{K}} \right]
             \left[ \frac{A_J}{A_{K}} \right]
             \left[ \frac{(K/L)_{J}}{(K/L)_{K}} \right]^\alpha
             \left[ \frac{h_{J}}{h_{K}} \right]^{1-\alpha} \\
             &=& \left[ \frac{.534}{.494} \right]
                 \left[ \frac{0.0538}{0.0511} \right]
                 \left[ \frac{180.2}{119.9} \right]^{1/3}
                 \left[ \frac{1801}{2434} \right]^{2/3}_{\phantom{\sum_i}} \\
      1.37   &=& (1.08)(1.37) (1.15) (0.82) .
\end{eqnarray*}
%
Where does the 37\% difference in GDP per person come from?
Most of the answer is capital: 
Japan has only an 37\% TFP advantage, 
and its higher capital per worker would (on its own) 
adds another 15\% to the difference.  
But hours worked, as we've seen, goes the other way:  
higher in Korea.  

Education also plays a role here.  
Since average years of school are higher in Korea, this will increase the difference
in TFP (what's left).  
One approach (mentioned in the notes) is to measure $H$ by $S$.  
Then a term $(H_J/H_K)^{2/3} = 0.913 $ will be added above.  
Or you could approximate $H$ by $ \exp(0.07 \times S)$, which would give you a somewhat different answer (0.937).  
Depending on your choice, the TFP difference will change, 
since TFP is (by construction) whatever's left.  


\item This part is up to you.  
One possibility is to look at measures of institutional quality 
and discuss the role you think they'll play in the future.
Institutions, on the whole, are good indicators of the future, 
since they tend not to change very quickly.  
\end{enumerate}
%\end{comment}

%\end{enumerate} \end{document} 