\documentclass[12pt]{exam}

\usepackage{ge05}
\usepackage{comment}
\usepackage{booktabs}
\usepackage[dvipdfm]{hyperref}
\urlstyle{rm}   % change fonts for url's (from Chad Jones)
\hypersetup{
    colorlinks=true,        % kills boxes
    allcolors=blue,
    pdfsubject={NYU Stern course GB 2303, Global Economy},
    pdfauthor={Dave Backus @ NYU},
    pdfstartview={FitH},
    pdfpagemode={UseNone},
%    pdfnewwindow=true,      % links in new window
%    linkcolor=blue,         % color of internal links
%    citecolor=blue,         % color of links to bibliography
%    filecolor=blue,         % color of file links
%    urlcolor=blue           % color of external links
% see:  http://www.tug.org/applications/hyperref/manual.html
}

% for ge05.sty
\def\ClassName{The Global Economy}
%\def\Category{Professor David Backus}
\def\Category{Backus \& Cooley}
\def\HeadName{Practice Problems B}
\newcommand{\phm}{\phantom{--}}
\newcommand{\NX}{\mbox{\it NX\/}}

\noprintanswers
\printanswers

\begin{document}
\parindent = 0.0in
\parskip = \bigskipamount
\thispagestyle{empty}%
\Head

\centerline{\large \bf \HeadName:  Labor Markets \& Trade}
\centerline{Revised:  \today}

\medskip
{\it This will not be collected or graded, 
but it's a good way to make sure you're up to speed.
We recommend you do it before the next class.}

\begin{questions}
% --------------------------------------------------------------------
\question As background to a consulting assignment, 
you have been asked to work through an example to illustrate
the impact of a minimum wage and payroll taxes on employment and unemployment.  
you start with these labor demand and supply functions:  
\begin{eqnarray*}
    \mbox{Demand:} && L^d(w) \;\;=\;\; 1/(1.5 w)^3 \\
    \mbox{Supply:} && L^s(w) \;\;=\;\; w^2 .
\end{eqnarray*}
%
\begin{parts}
\part Describe first how a frictionless labor market might work.
What wage rate clears the market (that is, equates supply and demand)?  
How much labor is employed?
What is the unemployment rate?  
Sketch the result in a supply/demand diagram.  

\part Now consider introducing a minimum wage.  If the minimum wage is 
$w_m = 1$, how much labor is employed? 
How much is supplied?  
What is the unemployment rate?  
Show how this works in your diagram.  

\part Suppose the minimum wage is $w_m = 1/2$.  
How much labor is employed?  Supplied?  
What is the unemployment rate?  

\end{parts}

% --------------------------------------------------------------------
\question Protecting sugar (30 points).
The US has a long-running tradition of protecting sugar producers, 
going back to 1789 when Treasury Secretary Alexander Hamilton helped
pass a tariff on sugar imports.  
(Tariffs at that time were the major source of federal government revenue
and needed to service the country's debt.) 
Currently the US restricts the quantity of sugar imports (quotas) and guarantees
prices to US producers that are well above world prices.  

\begin{parts}
\part Who benefits from these policies? 
\part Who loses?  
\part Economists believe that the cost of sugar protection outweigh
the benefits, yet these policies have been in place for 200 years. 
Why?  
\end{parts}

\begin{solution}
\begin{parts}
\part Producers of sugar or sugar substitutes benefit.  
Sugar cane and sugar beet farmers also benefit, as do corn growers, 
who produce the ubiquitous ``high-fructose corn syrup,''
a common substitute for sugar in American soft drinks and candy.  
The Iowa primary usually illustrates this connection in stark terms.

\part Anyone who consumes or produces a product that contains sugar is hurt.
People who use and eat substitutes are also hurt, since their prices have gotten
more expensive.  

\part Sad to say, economists do not make policy. 
Policy is made by people who wanted to get reelected.  
Sugar producers are a small group, so you might think their influence
is small, but they have a strong interest in such policies. 
Everyone else, a much larger group, generally loses from these policies, 
but their losses, while large in total, are small to them individually, 
hence unlikely to change their votes.  
\end{parts} 
\end{solution} 

% --------------------------------------------------------------------
\question Dumping coffee (30 points).
You are working for Illy USA, the American subsidiary of 
an Italian company that imports and sells coffee products in the US.
You boss tells you that domestic coffee roasters have filed a dumping complaint 
against the company, 
but does not know what that means.  
He asks:  
%
\begin{parts} 
\part What must they show to sustain a judgement of dumping against Illy?  

\part What damage could this do to Illy USA?  

\part Why has dumping become more common in the recent past?  
\end{parts} 

\begin{solution}
\begin{parts}
\part They must show two things.  First, that Illy has priced
coffee ``unfairly'' in the US.
For example, that Illy sells coffee in the US below its estimated cost of production 
or below what it charges in Italy.  
Second, that they have suffered ``injury'' as a result.  

\part It could result in duties (tariffs) imposed on Illy's imports, 
not to mention the endless litigation associated with 
many dumping cases.  

\part antidumping cases have risen in popularity has 
World Trade Organization (WTO) rules have eliminated other 
trade barriers.  
\end{parts}
%
Wikipedia has a nice summary:  \\
\centerline{\url{http://en.wikipedia.org/wiki/Dumping_(pricing_policy)}}
\end{solution}


\end{questions}

\vfill \centerline{\it \copyright \ \number\year \
NYU Stern School of Business}

\end{document}

