\documentclass[letterpaper,12pt]{article}

\RequirePackage{GE05}
% this inputs graphicx, too
\RequirePackage{comment}
\RequirePackage[hypertex]{hyperref}
%\hypersetup{dvips}
%\hypersetup{backref,colorlinks=false,urlcolor=cyan,linkcolor=blue}
%\hypersetup{pdfpagemode=None,pdfstartview=FitH}

\def\ClassName{The Global Economy}
\def\Category{Mini-Case}
\def\HeadName{Financing India's Public Sector}

\begin{document}
\parindent = 0.0in
\parskip = 0.6\bigskipamount
\thispagestyle{empty}%
\Head

\centerline{\large \bf \HeadName}%
\centerline{Revised:  \today}

\bigskip
During the 1980s and 1990s, 
India dismantled --- or at least weakened --- 
the extensive web of regulations 
that the government used to control (throttle?) the economy.
Largely as a result, it has now approached Chinese levels of growth, 
with real GDP increasing by 8-10\% a year between 2003 and 2007.
One sign of possible trouble on the horizon, however, is 
the government budget.
With an election scheduled for May,
the current government faces heavy political pressure 
to spend money.  

Your mission is to assess the situation:
the factors affecting the government budget 
and their likely impact on the economy.  
%
\begin{itemize}
\item What are the pressures on India's federal government budget?  
On spending?  On revenue?  
\item What is the projected budget deficit?  The primary deficit? 
\item How is the deficit financed?   
\item What is the ratio of public debt to GDP? 
How is it expected to change over the next two years?
\item What concerns might you have if you were
\begin{itemize}
\item An Indian business leader?  
\item A foreign investor in Indian companies?
\item A poor person in rural India?
\end{itemize}
\end{itemize}


Attached are excerpts from various EIU reports:
the Country Profile, the Country Report, and the Country Risk Service.
All of this material \copyright \ 2008 Economist Intelligence Unit. 
You can access the originals through NYU's 
\href{http://library.nyu.edu/vbl/}
{Virtual Business Library}.


%\vfill \centerline{\it \copyright \ \number\year \ NYU Stern School of Business}

\end{document} 

