\documentclass[letterpaper,12pt]{article}

\RequirePackage{GE05}
% this inputs graphicx, too
\RequirePackage{comment}
\RequirePackage[hypertex]{hyperref}
%\hypersetup{dvips}
%\hypersetup{backref,colorlinks=false,urlcolor=cyan,linkcolor=blue}
%\hypersetup{pdfpagemode=None,pdfstartview=FitH}

\def\ClassName{The Global Economy}
\def\Category{Mini-Case}
\def\HeadName{Second-Guessing Greenspan}

\begin{document}
\parindent = 0.0in
\parskip = 0.6\bigskipamount
\thispagestyle{empty}%
\Head

\centerline{\large \bf \HeadName}%
\centerline{Revised:  \today}

\bigskip 
Washington, February 2, 2000.   
As President of the Federal Reserve Bank of New York, 
you are a voting member of the Federal Open Market Committee, 
the body that makes decisions about US monetary policy.
Yesterday, the Committee heard from a number of experts 
about the state of the US economy (minutes attached).  
As you enter the new millennium, 
the dot-com boom continues, with employment and 
stock prices at all-time highs.
Is this unalloyed good news, 
or are there signs of trouble lurking in the data?  
Is the economy ``overheating"?  
You carefully weigh the evidence,  
working your way through a series of questions:   
%
\begin{itemize}
\item Economic conditions.   
What indicators are cited about the state of the US economy?  
What do they suggest about growth?  Inflation?  

\item Interpretation.  Does the evidence point to 
a shift in aggregate supply or demand?
How might you distinguish the two?  
Does it make a difference?  

\item Areas of concern.  Are there any sources of immediate concern?   
Unusual circumstances that demand further thought?  

\item Policy.  On the whole, do you think Greenspan's proposal for 
a 25 bp increase in the fed funds target is appropriate?  
If not, what change would you recommend?   

\end{itemize}

The attached material was adapted by Giulio Recchia (MBA 07)
from minutes posted by the Federal Reserve Board at:

\bigskip
\centerline{
\href{http://www.federalreserve.gov/FOMC/}{http://www.federalreserve.gov/FOMC/}
}

\vfill \centerline{\it \copyright \ \number\year \ NYU Stern
School of Business}

\end{document} 

