\documentclass[letterpaper,12pt]{article}

\RequirePackage{GE05}
% this inputs graphicx, too
\RequirePackage{comment}
\RequirePackage[hypertex]{hyperref}
%\hypersetup{dvips}
%\hypersetup{backref,colorlinks=false,urlcolor=cyan,linkcolor=blue}
%\hypersetup{pdfpagemode=None,pdfstartview=FitH}

\def\ClassName{The Global Economy}
\def\Category{Mini-Case}
\def\HeadName{Hyperinflation in Zimbabwe}

\begin{document}
\parindent = 0.0in
\parskip = 0.6\bigskipamount
\thispagestyle{empty}%
\Head

\centerline{\large \bf \HeadName}%
\centerline{Revised:  \today}

\bigskip 
Hyperinflations --- inflation of 100\% a year or more --- 
are a recurring feature of the economic landscape.
Famous examples include Germany in the 1920s and again after World War II, 
Bolivia and Israel in the 1980s, 
Bosnia-Herzegovina and Brazil in the 1990s, 
and Argentina several times.
Right now, Zimbabwe is the reigning champion, 
with an inflation rate well over 1000\%.  
Your mission is to explain why:  
%
\begin{itemize}
\item What is Zimbabwe's current economic situation?
Besides inflation, how is the economy performing?  
\item What are the roots of the current economic situation?
\item How would you describe the chain of events that led 
to such high inflation rates?  
What is the root cause?  The solution?  
\end{itemize}

Attached are excerpts from various EIU reports:
the Country Profile, the Country Report, and the Country Risk Service.
All of this material \copyright \ 2008 Economist Intelligence Unit. 
You can access the originals through NYU's 
\href{http://library.nyu.edu/vbl/}
{Virtual Business Library}.


%\vfill \centerline{\it \copyright \ \number\year \ NYU Stern
%School of Business}

\end{document} 

